---
title: "LaunchLens"
subtitle: "Student Startup Signals"
author: ""
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: false
    theme: null
    css: styles.css
    includes:
      after_body: site-footer.html
params:
  data_file: "data/global_startup_success_dataset-selected-columns.csv"
  seed: 2026
---

```{r setup, include=FALSE}
if (!exists("params", inherits = FALSE)) {
  params <- list(
    data_file = "data/global_startup_success_dataset-selected-columns.csv",
    seed = 2026
  )
}

knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, error = FALSE,
  fig.width = 9, fig.height = 5.5, dpi = 120
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(glmnet)
library(recipes)
library(scales)
library(broom)

launch_palette <- list(
  violet = "#310A31",
  mauve = "#847996",
  teal = "#88B7B5",
  sage = "#A7CAB1",
  cream = "#F4ECD6",
  background = "#17111F",
  surface = "#21182C",
  grid = "#44384F"
)

theme_set(
  theme_minimal(base_size = 12, base_family = "sans") +
    theme(
      plot.title.position = "plot",
      plot.background = element_rect(fill = launch_palette$surface, color = NA),
      panel.background = element_rect(fill = launch_palette$surface, color = NA),
      panel.grid.major = element_line(color = launch_palette$grid, linewidth = 0.35),
      panel.grid.minor = element_blank(),
      plot.title = element_text(color = launch_palette$cream, face = "bold", size = 15),
      plot.subtitle = element_text(color = launch_palette$sage),
      axis.title = element_text(color = launch_palette$cream),
      axis.text = element_text(color = launch_palette$cream),
      strip.background = element_rect(fill = launch_palette$violet, color = NA),
      strip.text = element_text(color = launch_palette$cream, face = "bold"),
      legend.position = "bottom",
      legend.background = element_rect(fill = launch_palette$surface, color = NA),
      legend.key = element_rect(fill = launch_palette$surface, color = NA),
      legend.text = element_text(color = launch_palette$cream),
      legend.title = element_text(color = launch_palette$cream)
    )
)

required_columns <- c(
  "Startup Name", "Founded Year", "Country", "Industry", "Funding Stage",
  "Total Funding ($M)", "Number of Employees", "Annual Revenue ($M)",
  "Valuation ($B)", "Success Score"
)

raw <- read.csv(
  params$data_file,
  check.names = FALSE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA", "N/A", "null")
)

# Treat whitespace-only cells as missing before checking for empty records.
raw[] <- lapply(raw, function(x) {
  if (is.character(x)) {
    x <- trimws(x)
    x[x == ""] <- NA_character_
  }
  x
})

missing_columns <- setdiff(required_columns, names(raw))
if (length(missing_columns) > 0) {
  stop("Missing required columns: ", paste(missing_columns, collapse = ", "))
}

input_rows <- nrow(raw)
# Remove rows that contain no information at all.
raw <- raw[rowSums(!is.na(raw)) > 0, , drop = FALSE]
has_data <- nrow(raw) > 0
```

<div class="site-nav">
<a class="brand" href="#">Launch<span>Lens</span></a>
<div class="nav-links">
<a href="#startup-roadmap">Start</a>
<a href="#problem-to-solution">Build</a>
<a href="#financial-foundations">Money</a>
<a href="#signal-lab">Tools</a>
<a href="#research-lab">Research</a>
</div>
</div>

<div class="hero">
<div class="hero-copy">
<div class="eyebrow">A practical startup guide for students</div>
<h1>Build the idea<br><span>people actually need.</span></h1>
<p>Learn how to find a real problem, test a solution, understand the numbers, and make smarter startup decisions—one experiment at a time.</p>
<div class="hero-actions">
<a class="button primary" href="#startup-roadmap">Start your roadmap</a>
<a class="button secondary" href="#signal-lab">Try the tools</a>
</div>
</div>
<div class="signal-card">
<div class="signal-orbit orbit-one"></div>
<div class="signal-orbit orbit-two"></div>
<div class="signal-core">LL</div>
<div class="signal-label label-one">funding</div>
<div class="signal-label label-two">traction</div>
<div class="signal-label label-three">context</div>
</div>
</div>

<div class="metric-strip">
  <div><strong>6</strong><span>startup stages</span></div>
  <div><strong>3</strong><span>student calculators</span></div>
  <div><strong>R</strong><span>evidence-backed research lab</span></div>
</div>

## Startup roadmap

<p class="section-intro">A startup is not just a small company or a clever app. It is a team searching for a repeatable way to solve a meaningful problem—and a sustainable way to deliver that solution.</p>

<div class="card-grid three">
<article class="learning-card">
<span class="card-kicker">01 · Problem</span>
<h3>Find a real struggle</h3>
<p>Start with a specific person, a repeated frustration, and the cost of leaving it unsolved.</p>
</article>
<article class="learning-card">
<span class="card-kicker">02 · Evidence</span>
<h3>Replace guesses with learning</h3>
<p>Talk to people, study current workarounds, and look for behavior—not polite compliments.</p>
</article>
<article class="learning-card">
<span class="card-kicker">03 · Experiment</span>
<h3>Test the smallest promise</h3>
<p>Use a prototype, landing page, pilot, or manual service before building the full product.</p>
</article>
</div>

<div class="journey">
<div class="journey-step"><span>1</span><strong>Problem</strong><small>Who struggles, and why?</small></div>
<div class="journey-step"><span>2</span><strong>Insight</strong><small>What does evidence reveal?</small></div>
<div class="journey-step"><span>3</span><strong>Solution</strong><small>What is your smallest promise?</small></div>
<div class="journey-step"><span>4</span><strong>Validation</strong><small>Will people act or pay?</small></div>
<div class="journey-step"><span>5</span><strong>Launch</strong><small>Can you deliver reliably?</small></div>
<div class="journey-step"><span>6</span><strong>Learn</strong><small>What should change next?</small></div>
</div>

## Problem to solution

<div class="split-panel">
<div>
<span class="card-kicker">Problem statement</span>
<h3>Make the problem narrow enough to test</h3>
<p class="formula">[Specific user] struggles to [make progress] when [situation] because [root cause], which leads to [measurable cost].</p>
<p>“Students need a productivity app” is broad. “First-year commuters miss campus events because information is scattered across five channels” is testable.</p>
</div>
<div>
<span class="card-kicker">Solution hypothesis</span>
<h3>Describe an outcome, not a feature pile</h3>
<p class="formula">We believe [solution] will help [user] achieve [outcome]. We will know when [behavior or metric] changes.</p>
<p>Your first solution is a hypothesis. Learning that it is wrong early is progress, not failure.</p>
</div>
</div>

<div class="checklist">
<h3>Quick problem check</h3>
<label><input type="checkbox"> I can name one specific user group.</label>
<label><input type="checkbox"> The problem happens repeatedly or has a high cost.</label>
<label><input type="checkbox"> People already use a workaround.</label>
<label><input type="checkbox"> I can reach these users for interviews or tests.</label>
<label><input type="checkbox"> I can observe a behavior that would prove improvement.</label>
</div>

## Ideas and brainstorming

<p class="section-intro">Strong startup ideas usually come from close observation: inconvenient workflows, excluded users, expensive delays, new technology, or a process you understand unusually well.</p>

<div class="card-grid four">
<article class="learning-card"><span class="card-kicker">Observe</span><h3>Keep a friction log</h3><p>Record repeated complaints, awkward workarounds, waiting, waste, and tasks people avoid.</p></article>
<article class="learning-card"><span class="card-kicker">Reframe</span><h3>Ask “How might we?”</h3><p>Turn a complaint into an open challenge without deciding the answer too early.</p></article>
<article class="learning-card"><span class="card-kicker">Expand</span><h3>Create 20 options</h3><p>Separate idea generation from judgment. Include manual, service, software, and community solutions.</p></article>
<article class="learning-card"><span class="card-kicker">Narrow</span><h3>Score the evidence</h3><p>Rank urgency, access to users, willingness to act, testability, and your unfair insight.</p></article>
</div>

<div class="prompt-box">
<strong>15-minute idea sprint</strong>
<p>Choose one community you know. List five repeated problems, three current workarounds for each, and one way to test demand this week for the strongest problem.</p>
</div>

## Validation and MVP

<div class="split-panel">
<div>
<span class="card-kicker">Customer discovery</span>
<h3>Interview the past, not the imaginary future</h3>
<ul class="mini-list">
<li>“Tell me about the last time this happened.”</li>
<li>“What did you do instead?”</li>
<li>“What did that cost in time, money, or stress?”</li>
<li>“Who decides whether to try or buy a solution?”</li>
</ul>
<p>Avoid pitching during discovery. The goal is to understand behavior and decision-making.</p>
</div>
<div>
<span class="card-kicker">Minimum viable product</span>
<h3>Build the smallest credible test</h3>
<ul class="mini-list">
<li>Clickable prototype for usability</li>
<li>Landing page or waitlist for demand</li>
<li>Manual concierge service for delivery</li>
<li>Paid pilot or preorder for willingness to pay</li>
</ul>
<p>An MVP is a learning instrument. Choose it based on the riskiest assumption—not the most exciting feature.</p>
</div>
</div>

<div class="resource-band">
<div><strong>Evidence before scale</strong><span>The NSF I-Corps program uses experiential customer discovery to reduce commercialization risk.</span></div>
<a href="https://www.nsf.gov/funding/initiatives/i-corps" target="_blank" rel="noopener">Explore NSF I-Corps ↗</a>
</div>

## Legal basics

<div class="legal-note"><strong>Educational guidance, not legal advice.</strong> Requirements depend on your country, state, industry, team, and customers. Use official sources and consult a qualified attorney or accountant for decisions with legal or tax consequences.</div>

<div class="card-grid three">
<article class="learning-card"><span class="card-kicker">Founders</span><h3>Put agreements in writing</h3><p>Clarify roles, ownership, vesting, decision rights, departures, and who owns work created for the startup.</p></article>
<article class="learning-card"><span class="card-kicker">Entity</span><h3>Choose structure deliberately</h3><p>Structure affects taxes, liability, paperwork, fundraising, and ownership. Do not copy another startup blindly.</p><a class="resource-link" href="https://www.sba.gov/business-guide/launch-your-business/choose-business-structure" target="_blank" rel="noopener">SBA structure guide ↗</a></article>
<article class="learning-card"><span class="card-kicker">Identity</span><h3>Check names and IP</h3><p>A domain or social handle does not establish trademark rights. Search similar marks and understand what your brand identifies.</p><a class="resource-link" href="https://www.uspto.gov/trademarks/basics" target="_blank" rel="noopener">USPTO trademark basics ↗</a></article>
<article class="learning-card"><span class="card-kicker">Registration</span><h3>Register in the right order</h3><p>Check state or local formation, licenses, permits, tax registration, and banking requirements. In the U.S., form an entity before requesting its EIN.</p><a class="resource-link" href="https://www.irs.gov/businesses/small-businesses-self-employed/get-an-employer-identification-number" target="_blank" rel="noopener">IRS EIN guidance ↗</a></article>
<article class="learning-card"><span class="card-kicker">Data</span><h3>Protect customer information</h3><p>Collect only what you need, secure access, explain usage clearly, and review rules for minors, health, education, payments, and location data.</p></article>
<article class="learning-card"><span class="card-kicker">People</span><h3>Know who works for you</h3><p>Employee and contractor classifications have different tax, wage, benefit, and IP consequences.</p></article>
</div>

## Financial foundations

<p class="section-intro">You do not need an MBA to understand startup finance. You do need a clear view of cash, costs, price, unit economics, and how long you can keep learning.</p>

<div class="card-grid four">
<article class="learning-card"><span class="card-kicker">Costs</span><h3>Separate one-time and monthly costs</h3><p>Include software, equipment, licenses, insurance, contractors, marketing, payment fees, and a realistic buffer.</p></article>
<article class="learning-card"><span class="card-kicker">Price</span><h3>Connect price to value</h3><p>Test willingness to pay. Revenue is stronger evidence than sign-ups, and margin matters more than revenue alone.</p></article>
<article class="learning-card"><span class="card-kicker">Cash</span><h3>Track burn and runway</h3><p>Net burn is monthly cash out minus cash in. Runway estimates how many months remain at the current burn.</p></article>
<article class="learning-card"><span class="card-kicker">Funding</span><h3>Match money to the milestone</h3><p>Bootstrapping, grants, competitions, loans, and equity each change risk, control, and expectations.</p></article>
</div>

<div class="resource-band">
<div><strong>Build a realistic cost map</strong><span>The SBA recommends identifying one-time and monthly expenses before launch, then using them for profit, break-even, and funding estimates.</span></div>
<a href="https://www.sba.gov/business-guide/plan-your-business/calculate-your-startup-costs" target="_blank" rel="noopener">SBA startup-cost guide ↗</a>
</div>

## Signal lab

<p class="section-intro">Use these calculators to turn startup activity into questions you can act on. Results stay in your browser and are educational estimates—not guarantees or investment advice.</p>

<div class="tool-grid">
<article class="calculator-card">
<span class="card-kicker">Customer health</span>
<h3>Retention rate</h3>
<p>Estimate the share of starting customers who remain at the end of a period.</p>
<label>Customers at start<input id="retention-start" type="number" min="1" step="1" placeholder="100"></label>
<label>New customers added<input id="retention-new" type="number" min="0" step="1" placeholder="20"></label>
<label>Customers at end<input id="retention-end" type="number" min="0" step="1" placeholder="95"></label>
<button id="retention-calculate" type="button">Calculate retention</button>
<div class="tool-result" id="retention-result" aria-live="polite"><strong>—</strong><span>Enter your customer counts</span></div>
<small>Formula: (ending customers − new customers) ÷ starting customers</small>
</article>

<article class="calculator-card">
<span class="card-kicker">Cash health</span>
<h3>Runway estimate</h3>
<p>Estimate how long your current cash lasts if monthly net burn stays constant.</p>
<label>Cash available (USD)<input id="runway-cash" type="number" min="0" step="100" placeholder="12000"></label>
<label>Monthly cash out (USD)<input id="runway-out" type="number" min="0" step="50" placeholder="2500"></label>
<label>Monthly cash in (USD)<input id="runway-in" type="number" min="0" step="50" placeholder="500"></label>
<button id="runway-calculate" type="button">Estimate runway</button>
<div class="tool-result" id="runway-result" aria-live="polite"><strong>—</strong><span>Enter cash and monthly flows</span></div>
<small>Runway changes when revenue, spending, or timing changes.</small>
</article>

<article class="calculator-card readiness-card">
<span class="card-kicker">Learning progress</span>
<h3>Startup readiness snapshot</h3>
<p>Rate the evidence behind your startup. This is a reflection tool—not a probability of success.</p>
<label>Problem evidence <output id="problem-value">3</output><input class="readiness-input" id="problem-score" type="range" min="1" max="5" value="3"></label>
<label>Customer learning <output id="customer-value">3</output><input class="readiness-input" id="customer-score" type="range" min="1" max="5" value="3"></label>
<label>Early traction <output id="traction-value">3</output><input class="readiness-input" id="traction-score" type="range" min="1" max="5" value="3"></label>
<label>Financial clarity <output id="finance-value">3</output><input class="readiness-input" id="finance-score" type="range" min="1" max="5" value="3"></label>
<label>Team execution <output id="team-value">3</output><input class="readiness-input" id="team-score" type="range" min="1" max="5" value="3"></label>
<div class="tool-result" id="readiness-result" aria-live="polite"><strong>50/100</strong><span>Validation stage</span></div>
</article>
</div>

<div class="prediction-note">
<strong>What can you responsibly estimate?</strong>
<p>With consistent historical data, models can estimate retention risk, demand, revenue ranges, churn, cash runway, or a defined success score. They cannot know whether a startup will “succeed overall” with certainty. Predictions inherit the assumptions, missing information, and bias in their training data.</p>
</div>

```{r prepare-data, include=FALSE}

if (has_data) {
  dat <- raw |>
    transmute(
      startup_name = trimws(`Startup Name`),
      founded_year = suppressWarnings(as.numeric(`Founded Year`)),
      country = trimws(Country),
      industry = trimws(Industry),
      funding_stage = trimws(`Funding Stage`),
      total_funding_m = suppressWarnings(as.numeric(`Total Funding ($M)`)),
      employees = suppressWarnings(as.numeric(`Number of Employees`)),
      annual_revenue_m = suppressWarnings(as.numeric(`Annual Revenue ($M)`)),
      valuation_b = suppressWarnings(as.numeric(`Valuation ($B)`)),
      success_score = suppressWarnings(as.numeric(`Success Score`))
    ) |>
    filter(!is.na(success_score)) |>
    mutate(
      startup_age = pmax(as.integer(format(Sys.Date(), "%Y")) - founded_year, 0),
      across(c(country, industry, funding_stage), ~ na_if(.x, "")),
      across(c(country, industry, funding_stage), as.factor),
      # Log transforms reduce the leverage of unusually large startups.
      log_funding = log1p(pmax(total_funding_m, 0)),
      log_employees = log1p(pmax(employees, 0)),
      log_revenue = log1p(pmax(annual_revenue_m, 0)),
      log_valuation = log1p(pmax(valuation_b, 0))
    )

  if (nrow(dat) < 20) {
    stop("At least 20 rows with a numeric Success Score are needed; found ", nrow(dat), ".")
  }
  if (sd(dat$success_score, na.rm = TRUE) == 0) {
    stop("Success Score has no variation, so a regression cannot be estimated.")
  }

  quality <- tibble(
    variable = names(dat),
    type = vapply(dat, function(x) class(x)[1], character(1)),
    missing_n = vapply(dat, function(x) sum(is.na(x)), integer(1)),
    missing_pct = missing_n / nrow(dat),
    unique_n = vapply(dat, function(x) length(unique(x[!is.na(x)])), integer(1))
  )

}
```

## Research lab

<div class="research-banner">
<div><span class="card-kicker">R + LASSO</span><h3>Test startup signals with data</h3></div>
<p>The learning guides above help students make decisions today. This research lab asks a narrower question: which observed startup characteristics are associated with a defined <code>Success Score</code>?</p>
</div>

Two models are prepared. The earlier-stage model uses age, country, industry,
funding stage, total funding, and employee count. The all-observed model adds
annual revenue and valuation. Because revenue and valuation may be consequences
of success, the second model is descriptive rather than causal.

```{r data-availability, echo=FALSE, results='asis'}
if (!has_data) {
  cat(
    paste0(
      "> **Research data needed:** `", params$data_file,
      "` contains the expected column names but no populated records. All ",
      input_rows, " rows in the supplied file are blank. The learning platform ",
      "and calculators work now; R graphs and model results will appear after the ",
      "file is replaced with populated data."
    )
  )
}
```

### Dataset overview

The dashboard begins with a basic data-quality check before showing any model
results. Missing values, sample size, and category coverage matter because a
clean-looking graph can still be misleading when the underlying sample is thin.

```{r quality-table}
if (has_data) {
  knitr::kable(
    quality,
    digits = 3,
    col.names = c("Variable", "Type", "Missing", "Missing %", "Unique values"),
    caption = paste("Data quality summary for", nrow(dat), "usable startups")
  )
}
```

```{r overview, fig.height=4.5}
if (has_data) {
  summary_stats <- dat |>
    summarise(
      startups = n(),
      countries = n_distinct(country, na.rm = TRUE),
      industries = n_distinct(industry, na.rm = TRUE),
      mean_success = mean(success_score, na.rm = TRUE),
      median_success = median(success_score, na.rm = TRUE),
      sd_success = sd(success_score, na.rm = TRUE)
    )
  knitr::kable(summary_stats, digits = 2)

  ggplot(dat, aes(success_score)) +
    geom_histogram(bins = 25, fill = launch_palette$teal, color = launch_palette$surface) +
    geom_vline(
      xintercept = mean(dat$success_score, na.rm = TRUE),
      color = launch_palette$sage, linewidth = 0.9, linetype = 2
    ) +
    labs(
      title = "Distribution of startup success scores",
      subtitle = "Dashed line is the sample mean",
      x = "Success score", y = "Number of startups"
    )
}
```

### Factor comparisons

Pearson correlation measures linear association; Spearman correlation also detects
monotonic relationships and is less sensitive to extreme values. Correlation does
not establish that a factor causes startup success.

```{r correlations}
if (has_data) {
  numeric_for_cor <- dat |>
    select(
      success_score, founded_year, startup_age, total_funding_m,
      employees, annual_revenue_m, valuation_b
    )

  target_correlations <- setdiff(names(numeric_for_cor), "success_score") |>
    lapply(function(v) {
      complete <- complete.cases(numeric_for_cor[[v]], numeric_for_cor$success_score)
      tibble(
        factor = v,
        complete_n = sum(complete),
        pearson = if (sum(complete) >= 3)
          cor(numeric_for_cor[[v]][complete],
              numeric_for_cor$success_score[complete], method = "pearson") else NA_real_,
        spearman = if (sum(complete) >= 3)
          cor(numeric_for_cor[[v]][complete],
              numeric_for_cor$success_score[complete], method = "spearman") else NA_real_
      )
    }) |>
    bind_rows() |>
    arrange(desc(abs(spearman)))

  knitr::kable(
    target_correlations, digits = 3,
    col.names = c("Factor", "Complete N", "Pearson r", "Spearman rho"),
    caption = "Bivariate associations with Success Score"
  )
}
```

```{r numeric-plots, fig.height=7}
if (has_data) {
  relationship_data <- dat |>
    select(
      success_score, `Startup age` = startup_age,
      `Total funding ($M)` = total_funding_m,
      Employees = employees,
      `Annual revenue ($M)` = annual_revenue_m,
      `Valuation ($B)` = valuation_b
    ) |>
    pivot_longer(-success_score, names_to = "factor", values_to = "value")

  ggplot(relationship_data, aes(value, success_score)) +
    geom_point(alpha = 0.5, color = launch_palette$teal) +
    geom_smooth(method = "lm", se = TRUE, color = launch_palette$sage, linewidth = 0.7) +
    facet_wrap(~ factor, scales = "free_x", ncol = 2) +
    labs(
      title = "Numeric factors and startup success",
      subtitle = "Lines show separate unadjusted linear relationships",
      x = NULL, y = "Success score"
    )
}
```

#### Country, industry, and funding-stage relationships

Group averages can be unstable for categories represented by few startups.
Accordingly, each plot shows the group sample size and only displays the 12
largest groups.

```{r grouped-relations, fig.height=10}
if (has_data) {
  group_summary <- function(data, variable) {
    data |>
      filter(!is.na(.data[[variable]])) |>
      group_by(group = .data[[variable]]) |>
      summarise(
        n = n(),
        mean_success = mean(success_score, na.rm = TRUE),
        se = sd(success_score, na.rm = TRUE) / sqrt(n),
        .groups = "drop"
      ) |>
      slice_max(n, n = 12, with_ties = FALSE) |>
      mutate(
        lower = mean_success - 1.96 * se,
        upper = mean_success + 1.96 * se,
        group = reorder(paste0(group, " (n=", n, ")"), mean_success)
      )
  }

  group_plot <- function(summary_data, title) {
    ggplot(summary_data, aes(mean_success, group)) +
      geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.15, color = launch_palette$mauve) +
      geom_point(size = 2.4, color = launch_palette$teal) +
      labs(title = title, x = "Mean success score (95% interval)", y = NULL)
  }

  print(group_plot(group_summary(dat, "funding_stage"), "Success by funding stage"))
  print(group_plot(group_summary(dat, "industry"), "Success by industry"))
  print(group_plot(group_summary(dat, "country"), "Success by country"))
}
```

### LASSO regression

The data are split 80/20 into training and test sets. All imputation, dummy
encoding, removal of zero-variance columns, and standardization are learned only
from the training data. Cross-validation selects the penalty. The conservative
`lambda.1se` value is used for the final model; it usually produces a smaller,
more stable set of predictors than the minimum-error penalty.

```{r lasso-functions}
if (has_data) {
  set.seed(params$seed)
  train_ids <- sample(seq_len(nrow(dat)), size = floor(0.80 * nrow(dat)))
  train <- dat[train_ids, , drop = FALSE]
  test <- dat[-train_ids, , drop = FALSE]

  run_lasso <- function(train, test, predictors, model_name) {
    model_formula <- reformulate(predictors, response = "success_score")

    rec <- recipe(model_formula, data = train) |>
      step_unknown(all_nominal_predictors(), new_level = "Unknown") |>
      step_novel(all_nominal_predictors(), new_level = "New") |>
      step_impute_median(all_numeric_predictors()) |>
      step_dummy(all_nominal_predictors(), one_hot = TRUE) |>
      step_zv(all_predictors()) |>
      step_normalize(all_numeric_predictors())

    trained_recipe <- prep(rec, training = train, retain = TRUE)
    train_baked <- bake(trained_recipe, new_data = train)
    test_baked <- bake(trained_recipe, new_data = test)

    x_train <- as.matrix(select(train_baked, -success_score))
    y_train <- train_baked$success_score
    x_test <- as.matrix(select(test_baked, -success_score))
    y_test <- test_baked$success_score

    set.seed(params$seed)
    folds <- min(10, max(3, floor(nrow(x_train) / 5)))
    cv_fit <- cv.glmnet(
      x_train, y_train,
      alpha = 1,
      family = "gaussian",
      nfolds = folds,
      standardize = FALSE,
      type.measure = "mse"
    )

    predictions <- as.numeric(
      predict(cv_fit, newx = x_test, s = "lambda.1se")
    )
    baseline <- rep(mean(y_train), length(y_test))
    test_sse <- sum((y_test - predictions)^2)
    test_sst <- sum((y_test - mean(y_test))^2)

    metrics <- tibble(
      model = model_name,
      test_n = length(y_test),
      selected_predictors =
        sum(as.matrix(coef(cv_fit, s = "lambda.1se"))[-1, 1] != 0),
      lambda = cv_fit$lambda.1se,
      rmse = sqrt(mean((y_test - predictions)^2)),
      mae = mean(abs(y_test - predictions)),
      r_squared = if (test_sst > 0) 1 - test_sse / test_sst else NA_real_,
      baseline_rmse = sqrt(mean((y_test - baseline)^2))
    )

    coefficient_matrix <- as.matrix(coef(cv_fit, s = "lambda.1se"))
    coefficients <- tibble(
      factor = rownames(coefficient_matrix),
      coefficient = as.numeric(coefficient_matrix[, 1])
    ) |>
      filter(factor != "(Intercept)", coefficient != 0) |>
      mutate(abs_coefficient = abs(coefficient)) |>
      arrange(desc(abs_coefficient))

    predictions_table <- tibble(
      actual = y_test,
      predicted = predictions,
      residual = y_test - predictions
    )

    list(
      recipe = trained_recipe,
      fit = cv_fit,
      metrics = metrics,
      coefficients = coefficients,
      predictions = predictions_table
    )
  }

  early_predictors <- c(
    "startup_age", "country", "industry", "funding_stage",
    "log_funding", "log_employees"
  )
  all_predictors <- c(
    early_predictors, "log_revenue", "log_valuation"
  )

  early_lasso <- run_lasso(train, test, early_predictors, "Earlier-stage factors")
  all_lasso <- run_lasso(train, test, all_predictors, "All observed factors")
}
```

#### Out-of-sample performance

```{r model-performance}
if (has_data) {
  performance <- bind_rows(early_lasso$metrics, all_lasso$metrics)
  knitr::kable(
    performance, digits = 3,
    col.names = c(
      "Model", "Test N", "Selected", "Lambda", "RMSE", "MAE",
      "Test R-squared", "Mean-only RMSE"
    )
  )
}
```

A useful model should have RMSE below the mean-only baseline on held-out data.
Negative test R-squared means the model generalized worse than simply predicting
the test-set mean.

#### Selected factors

Because inputs are standardized, coefficient magnitudes are broadly comparable.
A positive coefficient is associated with a higher predicted success score,
holding the other included variables constant; a zero coefficient was excluded
by LASSO. Category coefficients are relative to the omitted reference category.

```{r selected-coefficients, fig.height=8}
if (has_data) {
  selected <- bind_rows(
    mutate(early_lasso$coefficients, model = "Earlier-stage factors"),
    mutate(all_lasso$coefficients, model = "All observed factors")
  )

  if (nrow(selected) == 0) {
    cat("At the conservative lambda.1se penalty, LASSO selected no predictors.")
  } else {
    print(
      knitr::kable(
        selected |>
          group_by(model) |>
          slice_max(abs_coefficient, n = 20, with_ties = FALSE) |>
          ungroup() |>
          select(model, factor, coefficient),
        digits = 3,
        col.names = c("Model", "Selected factor", "Standardized coefficient")
      )
    )

    selected |>
      group_by(model) |>
      slice_max(abs_coefficient, n = 15, with_ties = FALSE) |>
      ungroup() |>
      mutate(
        factor = reorder(factor, coefficient),
        direction = if_else(coefficient > 0, "Positive", "Negative")
      ) |>
      ggplot(aes(coefficient, factor, fill = direction)) +
      geom_col() +
      facet_wrap(~ model, scales = "free_y") +
      scale_fill_manual(values = c(
        Positive = launch_palette$teal,
        Negative = launch_palette$mauve
      )) +
      labs(
        title = "Strongest factors retained by LASSO",
        subtitle = "Standardized coefficients at the one-standard-error penalty",
        x = "Coefficient", y = NULL, fill = "Association"
      )
  }
}
```

#### Prediction check

```{r prediction-check, fig.height=4.8}
if (has_data) {
  prediction_plot_data <- bind_rows(
    mutate(early_lasso$predictions, model = "Earlier-stage factors"),
    mutate(all_lasso$predictions, model = "All observed factors")
  )

  ggplot(prediction_plot_data, aes(actual, predicted)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, color = launch_palette$mauve) +
    geom_point(alpha = 0.7, color = launch_palette$teal) +
    facet_wrap(~ model) +
    coord_equal() +
    labs(
      title = "Held-out predictions versus actual scores",
      subtitle = "Points closer to the dashed 45-degree line are more accurate",
      x = "Actual success score", y = "Predicted success score"
    )
}
```

### Research conclusions

- These results identify **associations**, not causal effects. Funding, hiring,
  revenue, valuation, and success may influence one another.
- The startup name is excluded because it is an identifier, not a reusable
  explanatory factor.
- LASSO may choose one variable from a group of highly correlated variables and
  shrink the others to zero. A zero coefficient does not prove that a factor is
  irrelevant.
- Country and industry estimates need adequate representation. Results for small
  categories are unstable and should not drive decisions alone.
- The held-out test metrics are the best guide to whether the relationships
  generalize beyond this sample.
- A stronger causal analysis would require time-stamped predictors measured before
  the success outcome, a clear definition of `Success Score`, and preferably
  repeated observations over time.

### Research methodology

The analysis uses an 80/20 training-test split. Preprocessing is learned from the
training data only: missing numeric values are median-imputed, new and missing
categories are handled explicitly, categorical variables are dummy encoded, and
numeric predictors are standardized. Ten-fold cross-validation (or fewer folds
for small samples) selects the LASSO penalty. The conservative one-standard-error
penalty is reported to favor a simpler model.

The full R source is available in the project repository so every figure and
metric can be rebuilt when the dataset changes.

#### Reproducibility details

The report was generated with R and records its complete package environment in
the source build. This keeps the public page focused while preserving the
technical details in the repository.

```{r session-info, include=FALSE}
sessionInfo()
```
