---
title: "LaunchLens"
subtitle: "Student Startup Signals"
author: ""
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: false
    theme: null
    css: styles.css
params:
  data_file: "data/global_startup_success_dataset-selected-columns.csv"
  seed: 2026
---

```{r setup, include=FALSE}
if (!exists("params", inherits = FALSE)) {
  params <- list(
    data_file = "data/global_startup_success_dataset-selected-columns.csv",
    seed = 2026
  )
}

knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, error = FALSE,
  fig.width = 9, fig.height = 5.5, dpi = 120
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(glmnet)
library(recipes)
library(scales)
library(broom)

launch_palette <- list(
  violet = "#310A31",
  mauve = "#847996",
  teal = "#88B7B5",
  sage = "#A7CAB1",
  cream = "#F4ECD6",
  background = "#17111F",
  surface = "#21182C",
  grid = "#44384F"
)

theme_set(
  theme_minimal(base_size = 12, base_family = "sans") +
    theme(
      plot.title.position = "plot",
      plot.background = element_rect(fill = launch_palette$surface, color = NA),
      panel.background = element_rect(fill = launch_palette$surface, color = NA),
      panel.grid.major = element_line(color = launch_palette$grid, linewidth = 0.35),
      panel.grid.minor = element_blank(),
      plot.title = element_text(color = launch_palette$cream, face = "bold", size = 15),
      plot.subtitle = element_text(color = launch_palette$sage),
      axis.title = element_text(color = launch_palette$cream),
      axis.text = element_text(color = launch_palette$cream),
      strip.background = element_rect(fill = launch_palette$violet, color = NA),
      strip.text = element_text(color = launch_palette$cream, face = "bold"),
      legend.position = "bottom",
      legend.background = element_rect(fill = launch_palette$surface, color = NA),
      legend.key = element_rect(fill = launch_palette$surface, color = NA),
      legend.text = element_text(color = launch_palette$cream),
      legend.title = element_text(color = launch_palette$cream)
    )
)

required_columns <- c(
  "Startup Name", "Founded Year", "Country", "Industry", "Funding Stage",
  "Total Funding ($M)", "Number of Employees", "Annual Revenue ($M)",
  "Valuation ($B)", "Success Score"
)

raw <- read.csv(
  params$data_file,
  check.names = FALSE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA", "N/A", "null")
)

# Treat whitespace-only cells as missing before checking for empty records.
raw[] <- lapply(raw, function(x) {
  if (is.character(x)) {
    x <- trimws(x)
    x[x == ""] <- NA_character_
  }
  x
})

missing_columns <- setdiff(required_columns, names(raw))
if (length(missing_columns) > 0) {
  stop("Missing required columns: ", paste(missing_columns, collapse = ", "))
}

input_rows <- nrow(raw)
# Remove rows that contain no information at all.
raw <- raw[rowSums(!is.na(raw)) > 0, , drop = FALSE]
has_data <- nrow(raw) > 0
```

<div class="site-nav">
<a class="brand" href="#">Launch<span>Lens</span></a>
<div class="nav-links">
<a href="#dataset-overview">Data</a>
<a href="#factor-comparisons">Factors</a>
<a href="#lasso-regression">LASSO</a>
<a href="#conclusions">Conclusions</a>
</div>
</div>

<div class="hero">
<div class="hero-copy">
<div class="eyebrow">Student startup research · powered by R</div>
<h1>See the signals.<br><span>Build with context.</span></h1>
<p>LaunchLens turns startup data into clear, student-friendly evidence—so big ideas can begin with better questions.</p>
<div class="hero-actions">
<a class="button primary" href="#factor-comparisons">Explore the factors</a>
<a class="button secondary" href="#methodology">How it works</a>
</div>
</div>
<div class="signal-card">
<div class="signal-orbit orbit-one"></div>
<div class="signal-orbit orbit-two"></div>
<div class="signal-core">LL</div>
<div class="signal-label label-one">funding</div>
<div class="signal-label label-two">traction</div>
<div class="signal-label label-three">context</div>
</div>
</div>

<div class="metric-strip">
  <div><strong>2</strong><span>LASSO models</span></div>
  <div><strong>6</strong><span>research sections</span></div>
  <div><strong>R</strong><span>reproducible analysis</span></div>
</div>

## Introduction

This report asks which observed characteristics are associated with a startup's
`Success Score`. LASSO (least absolute shrinkage and selection operator) is useful
here because it can consider many numeric variables and category indicators at
once, shrink weak effects toward zero, and select a smaller set of useful
predictors.

Two specifications are used:

1. **Earlier-stage model:** age, country, industry, funding stage, total funding,
   and employee count.
2. **All-observed model:** the earlier-stage factors plus annual revenue and
   valuation.

The first is the more useful specification for early decision-making. Revenue and
valuation may be consequences of success, so the second model is descriptive and
should not be interpreted causally.

```{r data-availability, echo=FALSE, results='asis'}
if (!has_data) {
  cat(
    paste0(
      "> **Data issue:** `", params$data_file,
      "` contains the expected column names but no populated records. All ",
      input_rows, " rows in the supplied file are blank. The analysis below is fully ",
      "implemented and will run after the file is replaced with populated data."
    )
  )
}
```

```{r prepare-data, include=FALSE}

if (has_data) {
  dat <- raw |>
    transmute(
      startup_name = trimws(`Startup Name`),
      founded_year = suppressWarnings(as.numeric(`Founded Year`)),
      country = trimws(Country),
      industry = trimws(Industry),
      funding_stage = trimws(`Funding Stage`),
      total_funding_m = suppressWarnings(as.numeric(`Total Funding ($M)`)),
      employees = suppressWarnings(as.numeric(`Number of Employees`)),
      annual_revenue_m = suppressWarnings(as.numeric(`Annual Revenue ($M)`)),
      valuation_b = suppressWarnings(as.numeric(`Valuation ($B)`)),
      success_score = suppressWarnings(as.numeric(`Success Score`))
    ) |>
    filter(!is.na(success_score)) |>
    mutate(
      startup_age = pmax(as.integer(format(Sys.Date(), "%Y")) - founded_year, 0),
      across(c(country, industry, funding_stage), ~ na_if(.x, "")),
      across(c(country, industry, funding_stage), as.factor),
      # Log transforms reduce the leverage of unusually large startups.
      log_funding = log1p(pmax(total_funding_m, 0)),
      log_employees = log1p(pmax(employees, 0)),
      log_revenue = log1p(pmax(annual_revenue_m, 0)),
      log_valuation = log1p(pmax(valuation_b, 0))
    )

  if (nrow(dat) < 20) {
    stop("At least 20 rows with a numeric Success Score are needed; found ", nrow(dat), ".")
  }
  if (sd(dat$success_score, na.rm = TRUE) == 0) {
    stop("Success Score has no variation, so a regression cannot be estimated.")
  }

  quality <- tibble(
    variable = names(dat),
    type = vapply(dat, function(x) class(x)[1], character(1)),
    missing_n = vapply(dat, function(x) sum(is.na(x)), integer(1)),
    missing_pct = missing_n / nrow(dat),
    unique_n = vapply(dat, function(x) length(unique(x[!is.na(x)])), integer(1))
  )

}
```

## Dataset overview

The dashboard begins with a basic data-quality check before showing any model
results. Missing values, sample size, and category coverage matter because a
clean-looking graph can still be misleading when the underlying sample is thin.

```{r quality-table}
if (has_data) {
  knitr::kable(
    quality,
    digits = 3,
    col.names = c("Variable", "Type", "Missing", "Missing %", "Unique values"),
    caption = paste("Data quality summary for", nrow(dat), "usable startups")
  )
}
```

```{r overview, fig.height=4.5}
if (has_data) {
  summary_stats <- dat |>
    summarise(
      startups = n(),
      countries = n_distinct(country, na.rm = TRUE),
      industries = n_distinct(industry, na.rm = TRUE),
      mean_success = mean(success_score, na.rm = TRUE),
      median_success = median(success_score, na.rm = TRUE),
      sd_success = sd(success_score, na.rm = TRUE)
    )
  knitr::kable(summary_stats, digits = 2)

  ggplot(dat, aes(success_score)) +
    geom_histogram(bins = 25, fill = launch_palette$teal, color = launch_palette$surface) +
    geom_vline(
      xintercept = mean(dat$success_score, na.rm = TRUE),
      color = launch_palette$sage, linewidth = 0.9, linetype = 2
    ) +
    labs(
      title = "Distribution of startup success scores",
      subtitle = "Dashed line is the sample mean",
      x = "Success score", y = "Number of startups"
    )
}
```

## Factor comparisons

Pearson correlation measures linear association; Spearman correlation also detects
monotonic relationships and is less sensitive to extreme values. Correlation does
not establish that a factor causes startup success.

```{r correlations}
if (has_data) {
  numeric_for_cor <- dat |>
    select(
      success_score, founded_year, startup_age, total_funding_m,
      employees, annual_revenue_m, valuation_b
    )

  target_correlations <- setdiff(names(numeric_for_cor), "success_score") |>
    lapply(function(v) {
      complete <- complete.cases(numeric_for_cor[[v]], numeric_for_cor$success_score)
      tibble(
        factor = v,
        complete_n = sum(complete),
        pearson = if (sum(complete) >= 3)
          cor(numeric_for_cor[[v]][complete],
              numeric_for_cor$success_score[complete], method = "pearson") else NA_real_,
        spearman = if (sum(complete) >= 3)
          cor(numeric_for_cor[[v]][complete],
              numeric_for_cor$success_score[complete], method = "spearman") else NA_real_
      )
    }) |>
    bind_rows() |>
    arrange(desc(abs(spearman)))

  knitr::kable(
    target_correlations, digits = 3,
    col.names = c("Factor", "Complete N", "Pearson r", "Spearman rho"),
    caption = "Bivariate associations with Success Score"
  )
}
```

```{r numeric-plots, fig.height=7}
if (has_data) {
  relationship_data <- dat |>
    select(
      success_score, `Startup age` = startup_age,
      `Total funding ($M)` = total_funding_m,
      Employees = employees,
      `Annual revenue ($M)` = annual_revenue_m,
      `Valuation ($B)` = valuation_b
    ) |>
    pivot_longer(-success_score, names_to = "factor", values_to = "value")

  ggplot(relationship_data, aes(value, success_score)) +
    geom_point(alpha = 0.5, color = launch_palette$teal) +
    geom_smooth(method = "lm", se = TRUE, color = launch_palette$sage, linewidth = 0.7) +
    facet_wrap(~ factor, scales = "free_x", ncol = 2) +
    labs(
      title = "Numeric factors and startup success",
      subtitle = "Lines show separate unadjusted linear relationships",
      x = NULL, y = "Success score"
    )
}
```

### Country, industry, and funding-stage relationships

Group averages can be unstable for categories represented by few startups.
Accordingly, each plot shows the group sample size and only displays the 12
largest groups.

```{r grouped-relations, fig.height=10}
if (has_data) {
  group_summary <- function(data, variable) {
    data |>
      filter(!is.na(.data[[variable]])) |>
      group_by(group = .data[[variable]]) |>
      summarise(
        n = n(),
        mean_success = mean(success_score, na.rm = TRUE),
        se = sd(success_score, na.rm = TRUE) / sqrt(n),
        .groups = "drop"
      ) |>
      slice_max(n, n = 12, with_ties = FALSE) |>
      mutate(
        lower = mean_success - 1.96 * se,
        upper = mean_success + 1.96 * se,
        group = reorder(paste0(group, " (n=", n, ")"), mean_success)
      )
  }

  group_plot <- function(summary_data, title) {
    ggplot(summary_data, aes(mean_success, group)) +
      geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.15, color = launch_palette$mauve) +
      geom_point(size = 2.4, color = launch_palette$teal) +
      labs(title = title, x = "Mean success score (95% interval)", y = NULL)
  }

  print(group_plot(group_summary(dat, "funding_stage"), "Success by funding stage"))
  print(group_plot(group_summary(dat, "industry"), "Success by industry"))
  print(group_plot(group_summary(dat, "country"), "Success by country"))
}
```

## LASSO regression

The data are split 80/20 into training and test sets. All imputation, dummy
encoding, removal of zero-variance columns, and standardization are learned only
from the training data. Cross-validation selects the penalty. The conservative
`lambda.1se` value is used for the final model; it usually produces a smaller,
more stable set of predictors than the minimum-error penalty.

```{r lasso-functions}
if (has_data) {
  set.seed(params$seed)
  train_ids <- sample(seq_len(nrow(dat)), size = floor(0.80 * nrow(dat)))
  train <- dat[train_ids, , drop = FALSE]
  test <- dat[-train_ids, , drop = FALSE]

  run_lasso <- function(train, test, predictors, model_name) {
    model_formula <- reformulate(predictors, response = "success_score")

    rec <- recipe(model_formula, data = train) |>
      step_unknown(all_nominal_predictors(), new_level = "Unknown") |>
      step_novel(all_nominal_predictors(), new_level = "New") |>
      step_impute_median(all_numeric_predictors()) |>
      step_dummy(all_nominal_predictors(), one_hot = TRUE) |>
      step_zv(all_predictors()) |>
      step_normalize(all_numeric_predictors())

    trained_recipe <- prep(rec, training = train, retain = TRUE)
    train_baked <- bake(trained_recipe, new_data = train)
    test_baked <- bake(trained_recipe, new_data = test)

    x_train <- as.matrix(select(train_baked, -success_score))
    y_train <- train_baked$success_score
    x_test <- as.matrix(select(test_baked, -success_score))
    y_test <- test_baked$success_score

    set.seed(params$seed)
    folds <- min(10, max(3, floor(nrow(x_train) / 5)))
    cv_fit <- cv.glmnet(
      x_train, y_train,
      alpha = 1,
      family = "gaussian",
      nfolds = folds,
      standardize = FALSE,
      type.measure = "mse"
    )

    predictions <- as.numeric(
      predict(cv_fit, newx = x_test, s = "lambda.1se")
    )
    baseline <- rep(mean(y_train), length(y_test))
    test_sse <- sum((y_test - predictions)^2)
    test_sst <- sum((y_test - mean(y_test))^2)

    metrics <- tibble(
      model = model_name,
      test_n = length(y_test),
      selected_predictors =
        sum(as.matrix(coef(cv_fit, s = "lambda.1se"))[-1, 1] != 0),
      lambda = cv_fit$lambda.1se,
      rmse = sqrt(mean((y_test - predictions)^2)),
      mae = mean(abs(y_test - predictions)),
      r_squared = if (test_sst > 0) 1 - test_sse / test_sst else NA_real_,
      baseline_rmse = sqrt(mean((y_test - baseline)^2))
    )

    coefficient_matrix <- as.matrix(coef(cv_fit, s = "lambda.1se"))
    coefficients <- tibble(
      factor = rownames(coefficient_matrix),
      coefficient = as.numeric(coefficient_matrix[, 1])
    ) |>
      filter(factor != "(Intercept)", coefficient != 0) |>
      mutate(abs_coefficient = abs(coefficient)) |>
      arrange(desc(abs_coefficient))

    predictions_table <- tibble(
      actual = y_test,
      predicted = predictions,
      residual = y_test - predictions
    )

    list(
      recipe = trained_recipe,
      fit = cv_fit,
      metrics = metrics,
      coefficients = coefficients,
      predictions = predictions_table
    )
  }

  early_predictors <- c(
    "startup_age", "country", "industry", "funding_stage",
    "log_funding", "log_employees"
  )
  all_predictors <- c(
    early_predictors, "log_revenue", "log_valuation"
  )

  early_lasso <- run_lasso(train, test, early_predictors, "Earlier-stage factors")
  all_lasso <- run_lasso(train, test, all_predictors, "All observed factors")
}
```

### Out-of-sample performance

```{r model-performance}
if (has_data) {
  performance <- bind_rows(early_lasso$metrics, all_lasso$metrics)
  knitr::kable(
    performance, digits = 3,
    col.names = c(
      "Model", "Test N", "Selected", "Lambda", "RMSE", "MAE",
      "Test R-squared", "Mean-only RMSE"
    )
  )
}
```

A useful model should have RMSE below the mean-only baseline on held-out data.
Negative test R-squared means the model generalized worse than simply predicting
the test-set mean.

### Selected factors

Because inputs are standardized, coefficient magnitudes are broadly comparable.
A positive coefficient is associated with a higher predicted success score,
holding the other included variables constant; a zero coefficient was excluded
by LASSO. Category coefficients are relative to the omitted reference category.

```{r selected-coefficients, fig.height=8}
if (has_data) {
  selected <- bind_rows(
    mutate(early_lasso$coefficients, model = "Earlier-stage factors"),
    mutate(all_lasso$coefficients, model = "All observed factors")
  )

  if (nrow(selected) == 0) {
    cat("At the conservative lambda.1se penalty, LASSO selected no predictors.")
  } else {
    print(
      knitr::kable(
        selected |>
          group_by(model) |>
          slice_max(abs_coefficient, n = 20, with_ties = FALSE) |>
          ungroup() |>
          select(model, factor, coefficient),
        digits = 3,
        col.names = c("Model", "Selected factor", "Standardized coefficient")
      )
    )

    selected |>
      group_by(model) |>
      slice_max(abs_coefficient, n = 15, with_ties = FALSE) |>
      ungroup() |>
      mutate(
        factor = reorder(factor, coefficient),
        direction = if_else(coefficient > 0, "Positive", "Negative")
      ) |>
      ggplot(aes(coefficient, factor, fill = direction)) +
      geom_col() +
      facet_wrap(~ model, scales = "free_y") +
      scale_fill_manual(values = c(
        Positive = launch_palette$teal,
        Negative = launch_palette$mauve
      )) +
      labs(
        title = "Strongest factors retained by LASSO",
        subtitle = "Standardized coefficients at the one-standard-error penalty",
        x = "Coefficient", y = NULL, fill = "Association"
      )
  }
}
```

### Prediction check

```{r prediction-check, fig.height=4.8}
if (has_data) {
  prediction_plot_data <- bind_rows(
    mutate(early_lasso$predictions, model = "Earlier-stage factors"),
    mutate(all_lasso$predictions, model = "All observed factors")
  )

  ggplot(prediction_plot_data, aes(actual, predicted)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, color = launch_palette$mauve) +
    geom_point(alpha = 0.7, color = launch_palette$teal) +
    facet_wrap(~ model) +
    coord_equal() +
    labs(
      title = "Held-out predictions versus actual scores",
      subtitle = "Points closer to the dashed 45-degree line are more accurate",
      x = "Actual success score", y = "Predicted success score"
    )
}
```

## Conclusions

- These results identify **associations**, not causal effects. Funding, hiring,
  revenue, valuation, and success may influence one another.
- The startup name is excluded because it is an identifier, not a reusable
  explanatory factor.
- LASSO may choose one variable from a group of highly correlated variables and
  shrink the others to zero. A zero coefficient does not prove that a factor is
  irrelevant.
- Country and industry estimates need adequate representation. Results for small
  categories are unstable and should not drive decisions alone.
- The held-out test metrics are the best guide to whether the relationships
  generalize beyond this sample.
- A stronger causal analysis would require time-stamped predictors measured before
  the success outcome, a clear definition of `Success Score`, and preferably
  repeated observations over time.

## Methodology

The analysis uses an 80/20 training-test split. Preprocessing is learned from the
training data only: missing numeric values are median-imputed, new and missing
categories are handled explicitly, categorical variables are dummy encoded, and
numeric predictors are standardized. Ten-fold cross-validation (or fewer folds
for small samples) selects the LASSO penalty. The conservative one-standard-error
penalty is reported to favor a simpler model.

The full R source is available in the project repository so every figure and
metric can be rebuilt when the dataset changes.

### Reproducibility details

The report was generated with R and records its complete package environment in
the source build. This keeps the public page focused while preserving the
technical details in the repository.

```{r session-info, include=FALSE}
sessionInfo()
```
